manuscript_figures

### 2024-06-06 scripts/manuscript_figures.Rmd
updated stats for correlation analysis and made plot

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tigris)
library(car)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(ggplot2)
library(patchwork)
library(broom)
library(ggspatial)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)

@@

# Make Figure 2 and run statistical analysis
```{r}
# Aggregate the data by year, site, and transect and compute mean values
aggregated_data_cover <- merged.dat %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
            mean_kelp_cover = mean(COVER, na.rm = TRUE)) %>%
  ungroup()

@@

>>>>

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tigris)
library(car)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(ggspatial)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)

@@

# Make Figure 2 and run statistical analysis
```{r}
# Aggregate the data by year, site, and transect and compute mean values
aggregated_data_cover <- merged.dat.cover %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
            mean_kelp_cover = mean(COVER, na.rm = TRUE)) %>%
  ungroup()

@@

# Figure 5: Multiple regression (correlation analysis) between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
# combine all dataframes
```{r}
# left join can only combine two at a time
# combing kelp cover and fish biomass
master.dat1 <- left_join(kelp.filt, fish.filt)

#combine rugosity and kelp biomass
master.dat2 <- left_join(mean_rugosity, morph.filt)

# combine for final dataframe
master.dat <- left_join(master.dat1, master.dat2)

#delete unnecessary columns from the new joined data sheet 
all.params.data <- master.dat[, -c(1:7, 9:11, 14)]

#rename various columns in data frame
names(all.params.data)[1] <- "Kelp Cover (%)"
names(all.params.data)[2] <- "Fish Biomass (g/cm2)"
names(all.params.data)[3] <- "Rugosity"
names(all.params.data)[4] <- "Kelp Biomass (g/cm2)"

#reorder columns to show biomass by eaxchother
all.params.data <- all.params.data[, c(3, 1, 2, 4)]

view(all.params.data)

# Check for missing values
any_missing <- any(is.na(all.params.data))
if (any_missing) {
  # Handle missing values
  all.params.data <- na.omit(all.params.data)
}

# Compute the correlation matrix using Spearman correlation
cormat <- round(cor(all.params.data, method = "spearman"), 4)
head(cormat)

#melt the correlation matrix means it reassembles data frame to be more effective to complete corr matrix
#to long format
melted_cormat <- melt(cormat)
head(melted_cormat)

#visulaize the correlation matrix in general
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()

# Get lower and upper triangle of the correlation matrix
#Note that, a correlation matrix has redundant information. We’ll use the functions below to set half of it to NA
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}
# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}

#apply upper tri calculation to graphc
upper_tri <- get_upper_tri(cormat)
upper_tri

#melt the correlation matrix
#melt the correlation data and drop the rows with NA values 
melted_cormat <- melt(upper_tri, na.rm = TRUE)

#heatmap of correlation matrix
#negative correlations are in purple color and positive correlations in red
#scale_fill_gradient2 is used with the argument limit = c(-1,1) as correlation coefficients range from -1 to 1
#coord_fixed() : this function ensures that one unit on the x-axis is the same length as one unit on the y-axis
ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "midnightblue", high = "firebrick4", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Spearman's\nCorrelation") +
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))+
  coord_fixed()

#edit the number of sig figs listed in the value column
melted_cormat <- melted_cormat %>% mutate_at(vars(starts_with("value")), funs(round(., 2)))

# Create a ggheatmap with basic characteristics, etc. 
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "firebrick3", mid = "white", high = "dodgerblue3", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Spearman's\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))+
  coord_fixed()

# Print the heatmap
print(ggheatmap)

# Check for normality
shapiro.test(all.params.data$`Kelp Cover (%)`)
shapiro.test(all.params.data$`Fish Biomass (g/cm2)`) 
shapiro.test(all.params.data$Rugosity)
shapiro.test(all.params.data$`Kelp Biomass (g/cm2)`)

# data is not normal so move forward using aa spearmans rank correlation test

## Calculate p-values
p_values <- cor.mtest(all.params.data, method = "spearman")$p

## Convert the correlation matrix to a data frame
cormat_df <- reshape2::melt(cormat)

## Add p-values to the data frame
cormat_df$p_value <- p_values[cbind(cormat_df$Var1, cormat_df$Var2)]

## Create a new column to display stars based on p-values
cormat_df$sig <- apply_rules(cormat_df$p_value, rules = c(0.001, 0.01, 0.05), symbols = c("***", "**", "*"))

# Remove repeating rows
corузыdated_cormat_df <- cormat_df[!duplicated(cormat_df[, c("Var1", "Var2")]), ]

# Replace non-significant p-узыdues with "ns"
cormat_df$sig[cormat_df$p_value > 0.05] <- "ns"

# Print the modified data
print(cormat_df)

# Add correlation coefficients to the heatmap
ggheatmap +
  geom_text(data = cormat_df, aes(Var1, Var2, label = ifelse(Var1 == Var2, value, 
                                                             ifelse(sig == "", paste0(value, "\nns"), 
                                                             ifelse(duplicated(paste0(value, "\n", sig)), "", paste0(value, "\n", sig))))), 
            color = "black", size = 6) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 18, face="bold", color="black"),
        axis.text.y = element_text(size = 18, face="bold", color="black"),
        panel.grid.major = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        legend.justification = c(0.8, 0),
        legend.title = element_text(size = 18, face="bold", color="black"),
        legend.text = element_text(size = 20, face="bold", color="black"),
        legend.position = c(0.48, 0.75),
        legend.direction = "horizontal") +
  guides(fill = guide_colorbar(barwidth = 12, barheight = 2,
                               title.position = "top", title.hjust = 0.5, title.vjust = 1.0))

ggsave(filename = "output/figures/figure_5.png", device = "png", width = 10, height = 10)
```
